Temporal stability in human interaction networks: sector sizes, topological prominence and time activity

### 2015-10-05 paper.tex
rascunhada nova conclusao

\section{Conclusions}\label{sec:conc}
The most important result from the analysis of time evolution of the four email lists is certainly the time-independence observed not only for the activity but also for the properties of the networks themselves.
For example, the relative fractions of participants classified as hubs, intermediary and peripheral vertices remained practically constant along time across all email lists studied.
Furthermore, the PCA analysis of the topological metrics characterizing the networks also indicated that the contribution of each metric did not vary in time.
Centrality metrics were found to be the most relevant to characterize the network topology, followed by symmetry-related metrics, which were more relevant, with respect to variance, than clustering.

A systematic study of the activity of participants belonging to the three distinct Erd\"os sectors indicated simple patterns for hubs and peripheral vertices, while the network structure was governed by the intermediary vertices. These properties were shared by all email lists and were time-independent, being consistent with the literature. Moreover, both the distribution of Erd\"os sectors and the contribution from the metrics to the PCA were found to apply to networks from Facebook, Twitter and Participabr. We may therefore consider the classification of agents into Erd\"os sectors as a first step leading to a human typology which bridges exact sciences, with quantitative procedures for the classification, and human sciences, where there is a legacy in the observation of human types.

>>>>

\section{Conclusions}\label{sec:conc}
The temporal stability reported in this article is very counterintuitive.
These characteristics were expected from current scientific knowledge,
such as the fat tail structure and the preponderance of centrality measures,
but the very small standard deviations of principal components formation (Section~\ref{}),
the presence of the Erd\"os sectors even in networks with
few participants (Section~\ref{}),
and the recurrent activity patterns along different timescales (Section~\ref{}),
goes a step further in characterizing scale-free networks in the context
of the interaction of human individuals.
Maybe secondary, but nonetheless important, the importance of symmetry-related metrics,
which surpassed that of clustering coefficient,
with respect to dispersion of the system in the topological space,
might add to current understanding of how Digraphs differ from
undirected graphs.
Noteworthy is the also very stable fraction participants in each Erd\"os sectors
when the network reaches more than 200 participants (more that 1000 messages).
Benchmarks were derived from email list networks
and the analysis of
networks from Facebook,
Twitter and Participabr might ease hypothesizing
about the generality of such results.
Selected tables and figures sum dozens of pages
and are gathered in
the Supporting Information document.

Further work should expand the analysis to include
more types of networks and more metrics.
The data and software needed to reach these results
might also receive dedicated and in-depth
documentation as they enable a greater level of transparency
and work share,
which is adequate to both benchmarking
and the study of systems constituted
by human individuals (see Section~\ref{sec:data}).
The derived typology of hub, intermediary and peripheral participants
was used for semantic web and participatory democracy efforts,
and might be enhanced to yield scientific knowledge~\cite{pnud3,pnud4,pnud5}.
Also, we should pursue the publication of the visualization and sonification
processes used for this research~\cite{versinus,4hubs} and
of the extreme linguistic differences found in each Erd\"os sectors~\cite{rcText}.

% trabalhos de visualização (versinus), diferenciação de texto
% aplicação da tipologia para democracia participativa e dados ligados
% necessidade de publicar os dados em formatos ligados para comparacao da estrutura
% Future work on including more measures, other networks, sharing
% data in RDF for benchmarking, exploring larger timescales
% and of the data and OWL ontologies 
% 
%

%A systematic study of the activity of participants belonging to the three 
%distinct Erd\"os sectors indicated simple patterns for hubs and peripheral vertices,
%while the network structure was governed by the intermediary vertices.
%These properties were shared by all email lists and were time-independent,
%which is consistent with the literature.
%We may therefore consider the Erd\"os sectors as leading to a human typology which bridges exact sciences, with quantitative procedures for the classification, and human sciences, where there is a legacy in the observation of human types.
